---
title: "CannaQuant - Descriptives"
output:
  html_document: default
  pdf_document: default
date: "8 October 2024"
---

#   Data cleaned

in "clean data.R"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# for some basics, see https://www.statistikprofis.com/post/r-markdown

```

```{r, include = FALSE} 

# ESSENTIALS
rm(list=ls())
path <- paste0("data/")

## load libraries
library( data.table )
library( ggplot2 )
library( ggthemes )
library( tidyr )
library( stringr )

## themes and options
theme_set( theme_gdocs() )
options(scipen = 999)
green_shades_3 <- colorRampPalette(c("#DAF2D0", "#12501A"))(3)
green_shades_23 <- colorRampPalette(c("#DAF2D0", "#12501A"))(23)

# READ IN DATA
filename <- paste0(path,"all_countries_cleaned_data_",Sys.Date(),".rds")
input <- readRDS(file = filename)
dim(input) # 1530 * 19

data <- copy(input)

# PREPARE DATA



```

# Sample characteristics

First let's look at country-specific distributions

##   RESPONSES

How many valid responses are there in each country?

```{r echo = F}

#pdat <- data[, .N, by = iso]
pdat <- unique(data[,.(country,resp_total,resp_sexage,resp_sexageedu)])
pdat <- melt(pdat, id.vars = "country")
pdat$country_rev <- factor(pdat$country, levels = rev(levels(pdat$country)))
pdat$var <- factor(pdat$variable, labels = c("N total","N valid sex/age","N valid sex/age/edu"))
pdat$var_rev <- factor(pdat$var, levels = rev(levels(pdat$var)))
#pdat[, country := paste0(iso," (N=", N,")")]

ggplot(pdat) +
  geom_col(aes(x = country_rev, y = value, fill = var_rev), position = position_dodge()) +
  scale_fill_manual("", values = rev(green_shades_3)) + 
  scale_y_continuous("",breaks = seq(100,800,100)) + 
  guides(fill = guide_legend(reverse=TRUE)) +
  coord_flip()

rm(pdat)


```


### Lessons learned
***

1) In all countries (except CH): removing missing values in sex/age reduces the sample size substantially (more in PT than in other countries)

2) In CZ and ES: there was a survey programming problem resulting in many people not reporting their education


**for main analyses: either stick to the plan of complete education analyses or depart from study protocol and remove education as variable??**

##   SOCIODEMOGRAPHICS

How are sex and age distributed?

```{r echo = F}

temp <- copy(data)
temp <- temp[, .(share_men = sum(sex == "men")/.N,
                 age_mean = mean(age),
                 age_median = median(age),
                 age_IQR = paste0(quantile(age,0.25),"-",quantile(age,0.75))
                 #,
                 #edu_low = sum(edugroup == "low")/.N,
                 #edu_high = sum(edugroup == "high")/.N
                 ),
                 by = country]

print(temp)
rm(temp)

```


```{r echo = F}

pdat <- copy(data[edu.complete == T])
#levels(pdat$agegroup) <- rev(levels(pdat$agegroup))
pdat$country_rev <- factor(pdat$country, levels = rev(levels(pdat$country)))

ggplot(pdat, aes(x = educ, y = country_rev, fill= after_stat(x))) + 
  ggridges::geom_density_ridges_gradient(scale = 0.9, show.legend = F) + 
  scale_x_continuous("education (definition unclear - Arto?)", limits = c(0,5)) +
  scale_y_discrete("")

```

```{r include = F}
# age variations
summary(glm(age ~ relevel(country, ref = "Germany"), data, family = "gaussian"))

# education variations
summary(glm(educ ~ relevel(country, ref = "Germany"), pdat, family = "gaussian"))
rm(pdat)

```


### Lessons learned

AGE: Respondents are youngest in Czechia/Portugal and oldest in Germany/Sweden

EDUCATION: Pronounced variations in education across countries


## CANNABIS USE FREQUENCY

How is cannabis use frequency distributed in the different countries?

```{r echo = F}

pdat <- data[!is.na(freqgroup),.N, by = .(iso,country,freqgroup)]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := factor(paste0(iso," (N=", valid_N,")"))]
pdat$country_rev <- factor(pdat$country, levels = rev(levels(pdat$country)))
pdat$freqgroup_rev <- factor(pdat$freqgroup, levels = rev(levels(pdat$freqgroup)))

ggplot(pdat, aes(x = country_rev, y = prop, fill = freqgroup_rev)) +
  ggtitle("Number of use days in the past 30 days") +
  geom_col() +
  scale_fill_gdocs("") +
  scale_x_discrete("") +
  scale_y_continuous("", labels = scales::percent) + 
  guides(fill = guide_legend(reverse=TRUE)) +
  coord_flip()

data[!is.na(freqgroup), .(prev_daily = sum(freqgroup == "29-30")/.N), by = iso][order(prev_daily)]
rm(pdat)


```

### Lessons learned

Prevalence of (near) daily cannabis use varies across countries, with highest share in Sweden, Germany and Spain


## CAST SCORE

How is the CAST score distributed in the different countries?
Cutoff >= 7 indicates risk for severe use disorder

```{r echo = F}

pdat <- data[!is.na(cast_tot),.(iso,country,cast_tot)]

pdat[, ':=' (valid_N = .N), by = country]

pdat[, country := factor(paste0(iso,"\n(N=", valid_N,")"))]
pdat$country_rev <- factor(pdat$country, levels = rev(levels(pdat$country)))

pdat[, group := cast_tot >= 7]

#ggplot(pdat, aes(x = cast_tot, y = country_rev, fill= after_stat(x))) + 
#  ggridges::geom_density_ridges_gradient(scale = 0.9, show.legend = F) + 
#  geom_vline(xintercept = 6.5)
  
ggplot(pdat, aes(y = cast_tot, x = country_rev, fill = country_rev)) + 
  #geom_violin(alpha = 0.2, show.legend = F) +
  geom_jitter(alpha = 0.2, show.legend = F)+
  geom_boxplot(alpha = 0.2, show.legend = F) +
  geom_hline(yintercept = 7) +
  labs(y = "CAST Score (vertical line = cutoff 7)", x = "") +
  coord_flip()

data[, prop.table(table(highrisk))] # 54%
out <- data[,.(.N,cast_tot),by = country]
out <- out[,.(n = sum(cast_tot >=7)), by = .(country,N)]
print(out[,.(country,CAST_7_or_more = n/N, valid_N = N)])

rm(pdat)



```

### Lessons learned

1. About 54% are at risk for CUD (CAST>=7)
2. Highest percentages is highest in Sweden (72%) Spain (73%)



## THC CONTENT

How is the average THC content distributed in the different countries?

```{r echo = F}

pdat <- data[,.N, by = .(iso,country,thc_label)]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := factor(paste0(iso," (N=", valid_N,")"))]
pdat$country_rev <- factor(pdat$country, levels = rev(levels(pdat$country)))
pdat$thc_label_rev <- factor(pdat$thc_label, levels = rev(levels(pdat$thc_label)))

ggplot(pdat, aes(x = country_rev, y = prop, fill = thc_label_rev)) +
  ggtitle("THC content by country") +
  geom_col() +
  scale_fill_gdocs("") +
  scale_x_discrete("") +
  scale_y_continuous("", labels = scales::percent) + 
  guides(fill = guide_legend(reverse=TRUE)) +
  coord_flip()

rm(pdat)
```

### Lessons learned

1. Products <20% THC make up largest share 
2. Very low (>50%) THC knowledge in NL and PT - data problem?


# Cannabis amounts

## Product preference

```{r echo = F}

pdat <- data[,.N, by = .(iso,country,pref_type)]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := paste0(iso," (N=", valid_N,")")]
levels(pdat$pref_type) <- rev(levels(pdat$pref_type))

pdat <- data[,.N, by = .(iso,country,pref_type = factor(pref_type))]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := factor(paste0(iso," (N=", valid_N,")"))]
pdat$country_rev <- factor(pdat$country, levels = rev(levels(pdat$country)))
pdat$pref_type_rev <- factor(pdat$pref_type, levels = rev(levels(pdat$pref_type)))

ggplot(pdat, aes(x = country_rev, y = prop, fill = pref_type_rev)) +
  ggtitle("Product preference by country") +
  geom_col() +
  scale_fill_gdocs("") +
  scale_x_discrete("") +
  scale_y_continuous("", labels = scales::percent) + 
  guides(fill = guide_legend(reverse=TRUE)) +
  coord_flip()

rm(pdat)


```


### Lessons learned

1. High preference of cannabis flower in all countries
2. Resin plays a major role in ES, NL, PT and SE

## Crude Weight Assessment (CWA)

```{r include = F}

summary(data$cw_tot_flowers)
data[,quantile(cw_tot_flowers,0.95,na.rm = T), by = iso] # 


```

1. Removed top 5% -> very large amounts (CZ: >111g; DE: >70g; NL: 39g; PT: 153g; SK: >60g; ES: >150g; SE: >150g; CH: >104g) for illustration
2. Analysed only flowers due to small N of resin

```{r echo = F}

pdat <- data[complete.cases(cw_tot_flowers),.(iso,country,sex,agegroup,educ,freqgroup,cast_tot,cw_tot_flowers)]
pdat <- pdat[(iso == "CZ" & cw_tot_flowers <= 111) |
               (iso == "DE" & cw_tot_flowers <= 70) |
               (iso == "NL" & cw_tot_flowers <= 39) |
               (iso == "PT" & cw_tot_flowers <= 153) |
               (iso == "SK" & cw_tot_flowers <= 60) |
               (iso == "ES" & cw_tot_flowers <= 150) |
               (iso == "SE" & cw_tot_flowers <= 150) |
               (iso == "CH" &  cw_tot_flowers <= 104)]
pdat[, ':=' (valid_N =sum(.N)), by = country]
pdat[, country := paste0(iso,"\n(N=", valid_N,")")]

##  FREQUENCY
pdatfreq <- copy(pdat[complete.cases(freqgroup)])

#ggplot(pdatfreq) + 
#  geom_boxplot(aes(x = freqgroup, y = cw_tot_flowers, fill = freqgroup)) +
#  scale_fill_gdocs() +
#  facet_grid(country ~ .) +
#  labs(x = "Use Frequency",
#       y = "QWA: Grams of flowers in past 30 days")

ggplot(pdatfreq, aes(x = freqgroup, y = cw_tot_flowers, fill = freqgroup)) + 
  geom_boxplot(alpha = 0.6, show.legend = F) +
  geom_jitter(alpha = 0.2, show.legend = F) +
  scale_fill_gdocs() +
  labs(x = "Use Frequency",
       y = "CWA: Grams of flowers in past 30 days")

##  CAST RISK
pdatrisk <- copy(pdat[complete.cases(cast_tot)])
pdatrisk$cast_tot <- as.factor(pdatrisk$cast_tot)

ggplot(pdatrisk, aes(x = cast_tot, y = cw_tot_flowers, fill = cast_tot)) +
  geom_boxplot(alpha = 0.6, show.legend = F) +
  geom_jitter(alpha = 0.1, show.legend = F) +
  scale_fill_manual(values = green_shades_23) +
  labs(x = "CAST Score",
       y = "CWA: Grams of flowers in past 30 days")

rm(pdat, pdatfreq, pdatrisk)

```


### Lessons learned

1. Increased quantities (Crude Weight Assessment) among people with higher use frequency and higher CUD risk


## CannaQuant Assessment (CQ)

```{r include = F}

summary(data$total_cannabis)
data[,quantile(total_cannabis,0.95,na.rm = T), by = country] # 

```

1. Removed top 5% -> very large amounts (CZ: >75; DE: >75g; NL: >35g; PT: >170g; SK: >88g; ES: >117g; SE: >166g; CH: >193g) for illustration
2. Analysed only flowers due to small N of resin


```{r echo = F}

pdat <- data[complete.cases(total_flower),.(iso,country,sex,agegroup,educ,freqgroup,cast_tot,highrisk,total_flower)]
pdat <- pdat[(iso == "CZ" & total_flower <= 75) |
               (iso == "DE" & total_flower <= 75) |
               (iso == "NL" & total_flower <= 35) |
               (iso == "SK" & total_flower <= 88) |
               (iso == "ES" & total_flower <= 117) |
               (iso == "ES" & total_flower <= 166) |
               (iso == "CH" & total_flower <= 193)]
pdat[, ':=' (valid_N =sum(.N)), by = country]
pdat[, country := paste0(iso,"\n(N=", valid_N,")")]

##  FREQUENCY
pdatfreq <- copy(pdat[complete.cases(freqgroup)])

ggplot(pdatfreq, aes(x = freqgroup, y = total_flower, fill = freqgroup)) + 
  geom_boxplot(alpha = 0.6, show.legend = F) +
  geom_jitter(alpha = 0.2, show.legend = F) +
  scale_fill_gdocs() +
  labs(x = "Use Frequency",
       y = "CQ: Grams of flowers in past 30 days")

##  CAST RISK
pdatrisk <- copy(pdat[complete.cases(cast_tot)])
pdatrisk$cast_tot <- as.factor(pdatrisk$cast_tot)

ggplot(pdatrisk, aes(x = cast_tot, y = total_flower, fill = cast_tot)) +
  geom_boxplot(alpha = 0.6, show.legend = F) +
  geom_jitter(alpha = 0.1, show.legend = F) +
  scale_fill_manual(values = green_shades_23) +
  labs(x = "CAST Score",
       y = "CQ: Grams of flowers in past 30 days")


rm(pdat, pdatfreq, pdatrisk)

```


### Lessons learned

1. Increased quantities (CannaQuant) among people with higher use frequency and higher CUD risk

