---
title: "CannaQuant - Descriptives"
output: html_document
date: "2024-03-22"
---

#   Data cleaned

in "clean data.R"

Removed: missing value in sex OR age OR edu

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# for some basics, see https://www.statistikprofis.com/post/r-markdown

```

```{r, include = FALSE} 

# ESSENTIALS
rm(list=ls())
path <- paste0("data/")

## load libraries
library( data.table )
library( ggplot2 )
library( ggthemes )
library( tidyr )
library( stringr )

## themes and options
theme_set( theme_gdocs() )
options(scipen = 999)

# READ IN DATA

##  GERMANY
filename <- paste0(path,"DE_cleaned_data.csv")
input1 <- data.table(read.csv(filename))
dim(input1) # 370 * 14

##  SWITZERLAND
filename <- paste0(path,"DE_cleaned_data.csv")
input2 <- data.table(read.csv(filename))
dim(input2) # 370 * 14
input2$country <- "Switzerland"

# COMBINE DATA

data <- rbind(input1,
              input2)

# Define factors
data$sex <- factor(data$sex, levels = c("men","women"))
data$agegroup <- factor(data$agegroup, levels = c("18-24","25-34","35-44","45-54","55-73"))
data$edugroup <- factor(data$edugroup, levels = c("low","mid","high"))
data$freqgroup <- factor(data$freqgroup, levels = c("1-9","10-19","20-28","29-30"))
data$thc_label <- factor(data$thc_label, levels = c("0-4.9%","5-9.9%","10-14.9%","15-19.9%","20-24.9%","25-29.9%","30%+","unknown"))
data$highrisk <- factor(data$highrisk, labels = c("low","high"))


```

# Sample characteristics

First let's look at country-specific distributions

##   RESPONSES

How many valid responses are there in each country?

```{r echo = F}

pdat <- data[, .N, by = country]

ggplot(pdat) +
  geom_col(aes(x = country, y = N, fill = country), show.legend = F) +
  #geom_histogram(aes(fill = country, y = N), 
  #               position = "stack") +
  scale_fill_gdocs() + 
  scale_y_continuous(breaks = c(100,200,300,400))

rm(pdat)


```

##   SOCIODEMOGRAPHICS

How are sex, age, and education distributed?

```{r echo = F}

temp <- copy(data)
temp <- temp[, .(share_men = sum(sex == "men")/.N,
                 age_mean = mean(age),
                 age_median = median(age),
                 age_IQR = paste0(quantile(age,0.25),"-",quantile(age,0.75)),
                 edu_low = sum(edugroup == "low")/.N,
                 edu_high = sum(edugroup == "high")/.N),
                 by = country]

print(temp)
rm(temp)


```

```{r echo = F}

pdat <- copy(data)
levels(pdat$agegroup) <- rev(levels(pdat$agegroup))

ggplot(pdat, aes(x = agegroup, fill = edugroup)) +
  geom_histogram(stat = "count") +
  facet_grid(country ~ sex) +
  labs(x = "Age",
       y = "Number of respondents",
       fill = "Education") + 
  coord_flip()

rm(pdat)

```

## CANNABIS USE FREQUENCY

How is cannabis use frequency distributed in the different countries?

```{r echo = F}

pdat <- data[!is.na(freqgroup),.N, by = .(country,freqgroup)]
levels(pdat$freqgroup) <- rev(levels(pdat$freqgroup))

ggplot(pdat, aes(x = country, y = N, fill = freqgroup)) +
  geom_col() +
  scale_fill_gdocs() +
  labs(y = "Number of respondents")

out <- pdat[,.(valid_N =sum(N)),by = country]
print(out)
  
rm(pdat,out)

```

## CAST SCORE

How is the CAST score distributed in the different countries?
Cutoff >= 7 indicates risk for severe use disorder

```{r echo = F}

pdat <- data[!is.na(cast_tot),.N, by = .(country,cast_tot)]
pdat[, group := cast_tot >= 7]

ggplot(pdat, aes(x = cast_tot, y = N, fill = group)) +
  geom_col(show.legend = F) +
  geom_vline(xintercept = 6.5) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "CAST Score",
       y = "Number of respondents") 

out <- data[,.(.N,cast_tot),by = country]
out <- out[,.(n = sum(cast_tot >=7)), by = .(country,N)]
print(out[,.(country,CAST_7_or_more = n/N, valid_N = N)])

rm(pdat, out)

```

## THC CONTENT

How is the average THC content distributed in the different countries?

```{r echo = F}

pdat <- data[,.N, by = .(country,thc_label)]

ggplot(pdat, aes(x = country, y = N, fill = thc_label)) +
  geom_col() +
  scale_fill_gdocs(na.translate = T) +
  labs(y = "Number of respondents") 
  
out <- pdat[,.(valid_N =sum(N)),by = country]
print(out)

rm(pdat)

```

# Cannabis amounts

## Product preference

```{r echo = F}

pdat <- data[,.N, by = .(country,pref_type)]
levels(pdat$pref_type) <- rev(levels(pdat$pref_type))

ggplot(pdat, aes(x = country, y = N, fill = pref_type)) +
  geom_col(stat = "count") +
  scale_fill_gdocs() +
  labs(y = "Number of respondents")

out <- pdat[,.(valid_N =sum(N)),by = country]
print(out)
  
rm(pdat, out)


```

## Crude Weight Assessment (CWA)

```{r include = F}

summary(data$cw_tot_flowers)
data[,quantile(cw_tot_flowers,0.95,na.rm = T)] # 75g


```

- Removed very large amounts (more than 75g; 5% of responses) for illustration
- Analysed only flowers due to small N of resin

```{r echo = F}

pdat <- data[complete.cases(cw_tot_flowers),.(country,sex,agegroup,edugroup,freqgroup,highrisk,cw_tot_flowers)]
pdat <- pdat[cw_tot_flowers <= 75]


##  FREQUENCY
pdatfreq <- copy(pdat[complete.cases(freqgroup)])

ggplot(pdatfreq) + 
  geom_boxplot(aes(x = freqgroup, y = cw_tot_flowers, fill = freqgroup)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "Use Frequency",
       y = "QWA: Grams of flowers in past 30 days")

out <- pdatfreq[,.(valid_N =.N),by = country]
print(out)

##  CAST RISK
pdatrisk <- copy(pdat[complete.cases(highrisk)])

ggplot(pdat) + 
  geom_boxplot(aes(x = highrisk, y = cw_tot_flowers, fill = highrisk)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "CUD RISK (CAST >= 7)",
       y = "QWA: Grams of flowers in past 30 days")

out <- pdatrisk[,.(valid_N =.N),by = country]
print(out)

rm(pdat, pdatfreq, pdatrisk, out)

```

## CannaQuant Assessment (CQ)

```{r include = F}

summary(data$total_cannabis)
data[,quantile(total_cannabis,0.95,na.rm = T)] # 84g

```

- Removed very large amounts (more than 84g; 5% of responses) for illustration
- Analysed only flowers due to small N of resin


```{r echo = F}

pdat <- data[complete.cases(total_flower),.(country,sex,agegroup,edugroup,freqgroup,highrisk,total_flower)]
pdat <- pdat[total_flower <= 84]

##  FREQUENCY
pdatfreq <- copy(pdat[complete.cases(freqgroup)])

ggplot(pdatfreq) + 
  geom_boxplot(aes(x = freqgroup, y = total_flower, fill = freqgroup)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "Use Frequency",
       y = "CQ: Grams of flowers in past 30 days")

out <- pdatfreq[,.(valid_N =.N),by = country]
print(out)

##  CAST RISK
pdatrisk <- copy(pdat[complete.cases(highrisk)])

ggplot(pdat) + 
  geom_boxplot(aes(x = highrisk, y = total_flower, fill = highrisk)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "CUD RISK (CAST >= 7)",
       y = "CQ: Grams of flowers in past 30 days")

out <- pdatrisk[,.(valid_N =.N),by = country]
print(out)

rm(pdat, pdatfreq, pdatrisk, out)

```

```

