---
title: "CannaQuant - Descriptives"
output:
  html_document: default
  pdf_document: default
date: "28 July 2024"
---

#   Data cleaned

in "clean data.R"

Switzerland and Germany: missing value in sex OR age OR education

All other countries: missing value in sex OR age

Reason: many missing values in education in most countries --> Arto to check the data!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# for some basics, see https://www.statistikprofis.com/post/r-markdown

```

```{r, include = FALSE} 

# ESSENTIALS
rm(list=ls())
path <- paste0("data/")

## load libraries
library( data.table )
library( ggplot2 )
library( ggthemes )
library( tidyr )
library( stringr )

## themes and options
theme_set( theme_gdocs() )
options(scipen = 999)

# READ IN DATA
filename <- paste0(path,"all_countries_cleaned_data_",Sys.Date(),".rds")
input <- readRDS(file = filename)
dim(input) # 1405 * 18

data <- copy(input)

# PREPARE DATA



```

# Sample characteristics

First let's look at country-specific distributions

##   RESPONSES

How many valid responses are there in each country?

```{r echo = F}

pdat <- data[, .N, by = iso]
pdat[, country := paste0(iso," (N=", N,")")]

ggplot(pdat) +
  geom_col(aes(x = country, y = N, fill = country), show.legend = F) +
  scale_fill_gdocs() + 
  scale_y_continuous(breaks = c(100,200,300,400))

rm(pdat)


```

##   SOCIODEMOGRAPHICS

How are sex, age, and education distributed?

```{r echo = F}

temp <- copy(data)
temp <- temp[, .(share_men = sum(sex == "men")/.N,
                 age_mean = mean(age),
                 age_median = median(age),
                 age_IQR = paste0(quantile(age,0.25),"-",quantile(age,0.75)),
                 edu_low = sum(edugroup == "low")/.N,
                 edu_high = sum(edugroup == "high")/.N),
                 by = country]

print(temp)
rm(temp)

summary(glm(age ~ relevel(country, ref = "Germany"), data, family = "gaussian"))
print("Respondents are youngest in Czechia and oldest in Germany")


```

IGNORE THE EDUCATION DISTRIBUTION FOR NOW!

```{r echo = F}

pdat <- copy(data)
#levels(pdat$agegroup) <- rev(levels(pdat$agegroup))

ggplot(pdat, aes(x = agegroup, fill = edugroup)) +
  geom_histogram(stat = "count") +
  facet_grid(iso ~ sex) +
  labs(x = "Age",
       y = "Number of respondents",
       fill = "Education") + 
  coord_flip()

rm(pdat)

```

## CANNABIS USE FREQUENCY

How is cannabis use frequency distributed in the different countries?

```{r echo = F}

pdat <- data[!is.na(freqgroup),.N, by = .(iso,country,freqgroup)]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := paste0(iso," (N=", valid_N,")")]

#levels(pdat$freqgroup) <- rev(levels(pdat$freqgroup))

ggplot(pdat, aes(x = country, y = prop, fill = freqgroup)) +
  ggtitle("Number of use days in the past 30 days") +
  geom_col() +
  scale_fill_gdocs("") +
  scale_y_continuous("", labels = scales::percent)

rm(pdat)

print("About 50% of respondents use cannabis (near) daily in CH, DE and ES and the percentage is at about 25% in CZ and SK")


```

## CAST SCORE

How is the CAST score distributed in the different countries?
Cutoff >= 7 indicates risk for severe use disorder

```{r echo = F}

pdat <- data[!is.na(cast_tot),.N, by = .(iso,country,cast_tot)]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := paste0(iso,"\n(N=", valid_N,")")]

pdat[, group := cast_tot >= 7]

ggplot(pdat, aes(x = cast_tot, y = prop, fill = group)) +
  ggtitle("% of respondents with CAST sum score x") +
  geom_col(show.legend = F) +
  geom_vline(xintercept = 6.5) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  scale_y_continuous("", labels = scales::percent) +
  labs(x = "CAST Score") 

out <- data[,.(.N,cast_tot),by = country]
out <- out[,.(n = sum(cast_tot >=7)), by = .(country,N)]
print(out[,.(country,CAST_7_or_more = n/N, valid_N = N)])

rm(pdat)

print("About 50% are at risk for CUD but this percentage is higher in Spain (73%)")

```

## THC CONTENT

How is the average THC content distributed in the different countries?

```{r echo = F}

pdat <- data[,.N, by = .(iso,country,thc_label)]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := paste0(iso," (N=", valid_N,")")]

ggplot(pdat, aes(x = country, y = prop, fill = thc_label)) +
  geom_col() +
  scale_fill_gdocs(na.translate = T) +
  scale_y_continuous("", labels = scales::percent) +
  labs(y = "Number of respondents") 
  
rm(pdat)

print("Insufficient THC knowledge in CZ, ES, SK - data problem?")

```

# Cannabis amounts

## Product preference

```{r echo = F}

pdat <- data[,.N, by = .(iso,country,pref_type)]
pdat[, ':=' (valid_N =sum(N),
        prop = N/sum(N)), by = country]
pdat[, country := paste0(iso," (N=", valid_N,")")]
levels(pdat$pref_type) <- rev(levels(pdat$pref_type))

ggplot(pdat, aes(x = country, y = prop, fill = pref_type)) +
  geom_col(stat = "count") +
  scale_fill_gdocs() +
  labs(y = "Number of respondents")

rm(pdat)

print("High preference of cannabis flower in all countries; resin plays a role only in CH and ES")

```

## Crude Weight Assessment (CWA)

```{r include = F}

summary(data$cw_tot_flowers)
data[,quantile(cw_tot_flowers,0.95,na.rm = T), by = iso] # DE: 74g // CH: 104g


```

- Removed top 5% -> very large amounts (CZ: >111g; DE: >70g; SV: >60g; ES: >150g; CH: >104g) for illustration
- Analysed only flowers due to small N of resin

```{r echo = F}

pdat <- data[complete.cases(cw_tot_flowers),.(iso,country,sex,agegroup,edugroup,freqgroup,highrisk,cw_tot_flowers)]
pdat <- pdat[(iso == "CZ" & cw_tot_flowers <= 111) |
               (iso == "DE" & cw_tot_flowers <= 70) |
               (iso == "SK" & cw_tot_flowers <= 60) |
               (iso == "ES" & cw_tot_flowers <= 150) |
               (iso == "CH" &  cw_tot_flowers <= 104)]
pdat[, ':=' (valid_N =sum(.N)), by = country]
pdat[, country := paste0(iso,"\n(N=", valid_N,")")]

##  FREQUENCY
pdatfreq <- copy(pdat[complete.cases(freqgroup)])

ggplot(pdatfreq) + 
  geom_boxplot(aes(x = freqgroup, y = cw_tot_flowers, fill = freqgroup)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "Use Frequency",
       y = "QWA: Grams of flowers in past 30 days")

##  CAST RISK
pdatrisk <- copy(pdat[complete.cases(highrisk)])

ggplot(pdat) + 
  geom_boxplot(aes(x = highrisk, y = cw_tot_flowers, fill = highrisk)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "CUD RISK (CAST >= 7)",
       y = "QWA: Grams of flowers in past 30 days")

rm(pdat, pdatfreq, pdatrisk)

```

## CannaQuant Assessment (CQ)

```{r include = F}

summary(data$total_cannabis)
data[,quantile(total_cannabis,0.95,na.rm = T), by = country] # DE: 82g // CH: 193g

```

- Removed top 5% -> very large amounts (CZ: >65; DE: >75g; SK: >96g; ES: >138g; CH: >193g) for illustration
- Analysed only flowers due to small N of resin


```{r echo = F}

pdat <- data[complete.cases(total_flower),.(iso,country,sex,agegroup,edugroup,freqgroup,highrisk,total_flower)]
pdat <- pdat[(iso == "CZ" & total_flower <= 65) |
               (iso == "DE" & total_flower <= 75) |
               (iso == "SK" & total_flower <= 96) |
               (iso == "ES" & total_flower <= 138) |
               (iso == "CH" & total_flower <= 193)]
pdat[, ':=' (valid_N =sum(.N)), by = country]
pdat[, country := paste0(iso,"\n(N=", valid_N,")")]

##  FREQUENCY
pdatfreq <- copy(pdat[complete.cases(freqgroup)])

ggplot(pdatfreq) + 
  geom_boxplot(aes(x = freqgroup, y = total_flower, fill = freqgroup)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "Use Frequency",
       y = "CQ: Grams of flowers in past 30 days")

##  CAST RISK
pdatrisk <- copy(pdat[complete.cases(highrisk)])

ggplot(pdat) + 
  geom_boxplot(aes(x = highrisk, y = total_flower, fill = highrisk)) +
  scale_fill_gdocs() +
  facet_grid(country ~ .) +
  labs(x = "CUD RISK (CAST >= 7)",
       y = "CQ: Grams of flowers in past 30 days")

rm(pdat, pdatfreq, pdatrisk)

```

